---
title: "Introduction to R"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  pdf_document:
    df_print: kable
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Set Up

Use `install.packages("package_name")` to install packages from CRAN and `library(package_name)` to load packages for use in your current session.

```{r}
# Installing a package 
# install.packages("tidyverse")

# Loading the package for use
# library(tidyverse)
pacman::p_load(tidyverse)
```

# Data Exploration

-   **Dataset Structure**

```{r}
# Data structure
diamonds %>% glimpse()
```

-   **Top Rows in the Dataset**

```{r}
# top 6 rows
diamonds %>% head()
```

-   **Distribution of Categorical variables**

```{r}
# How many cut types are there
diamonds %>% count(cut)
```

-   **Dataset Summary Statistics**

```{r}
# Summary Statistics
diamonds %>% summary()
```

# Data Visualization

-   **Most tidyverse functions take a dataset as their first argument**

## Scatter Plot

```{r message=FALSE}
# Example of a Scatter Plot 
diamonds %>%                                                                    # the data
  sample_n(size = 1000) %>%                                                     # Sample a few rows
  ggplot(aes(x = carat, y = price, color = "lightsalmon"))+                     # Aesthetics mapping 
  geom_point(show.legend = F, alpha = .5)+                                      # Geom layer (Adding Scatter points)
  geom_smooth(method = "lm")+                                                   # Another geom layer (Add a smooth line)
  labs(title =  "Relationship between Price and Carat in the Diamonds Dataset")+# Adding plot title
  theme(legend.position = "none")+                                              # Remove unnecessary Legend
  theme_bw()                                                                    # Add a theme
```

## Box Plot

```{r}
# Example of a Box plot
diamonds %>%                                                             # the data
  ggplot(aes(x = cut, y = price, fill = cut))+                           # Aesthetics mapping 
  geom_boxplot(show.legend = F)+                                         # geom layer
  labs(title =  "Distribution of Price by Cut in the Diamonds Dataset")+ # adding plot title
  theme_bw()                                                             # Theme for the plot
```

## Histogram

```{r}
# Example of Histogram
diamonds %>%  
  ggplot(aes(x = price))+                                 
  geom_histogram(fill = "steelblue", bins = 30)+
  labs(title =  "Distribution of Price in the Diamonds Dataset")+ 
  theme_bw()
```

## Line Graphs for Time Series Data

```{r}
# Plotting time series data
library(timetk)

m4_daily %>%                     # the data set 
  filter(id == "D10") %>%        # filter for a specific group
  plot_time_series(              # Plotting function; two major arguments (date and value)
    .date_var = date, .value = value, .x_lab = "Date", .y_lab = "Value",.interactive = F
    )
```

## Anomaly detection

```{r message=FALSE}
# Checking Anomalies in data
m4_daily %>%                         # the data set 
  filter(id == "D10") %>%            # filter for a specific group
  plot_anomaly_diagnostics(          # Plotting function; two major Arguments (date and value)
    .date_var = date, .value = value, .x_lab = "Date", .y_lab = "Value",
    .interactive = F, .legend_show = F, .title = "Anomalies Highlighted in Red",
  )
```

## Visualizing Seasonality features in time series data

```{r fig.height=10, message=FALSE}
m4_daily %>%                    # the data set 
  filter(id == "D10") %>%       # filter for a specific group
  plot_seasonal_diagnostics(    # Plotting function; takes two major Arguments (date and value)
    .date_var = date, .value = value, .x_lab = "Date", .y_lab = "Value",
    .interactive = F, .title = "Seasonality breakdown by Weekday, Week, Month and Quarter"
  )
```
